Clicks and Cliques

### 2015-09-28 papersoul.Rnw
final version

% All chunks that make models or data procesing, at the begining
<<setup, include=FALSE, cache=FALSE,warning=FALSE,message=FALSE>>=
options(replace.assign=TRUE,width=50)
opts_chunk$set(fig.path='figure/graphics-', cache.path='cache/graphics-', fig.align='center', fig.width=8, fig.height=5, fig.show='hold', cache=TRUE, par=TRUE)
knit_hooks$set(par=function(before, options, envir){
if (before && options$fig.show!='none') par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
}, crop=hook_pdfcrop)

# packages
library(reshape2)
library(productplots) 
library(ggplot2)
library(stats)
library(randomForest)
library(maps)
library(fpc)
library(xtable)
library(car)
library(plyr)
library(GGally)
library(survey)
library(reldist)

@@

# Compute comunity proportion of attached people
lev <- levels(soul$QSB)
census.data %<>%
        mutate(QSB = factor(census.data$QSB, labels=lev) ) %>%
        left_join(att.pr[, c(1,2,4)], by = "QSB") %>% 
        mutate(QSB = reorder(QSB, CCEGRP2Attached))

@@

# Variable sets 
orig <- c(4:6, 8:61 ,148:152,165:175, 181:206, 210:213)
demo <- c(7,62:79,153:154,158:160,206:209,216,229)
reco <- c(80:120,141:147,161,177:179,217:228)
cons <- c(121:135,155:157, 162:164,176,180)

@@

# Compute attachment proportion for ordering QSB levels
wt     <- with(subset(soul,!is.na(WEIGHT)), WEIGHT/3)
des.wt <- svydesign(ids=~1, strata=~QSB, weights=~wt, data=subset(soul, !is.na(WEIGHT) ) )
att.pr <- svyby(~CCEGRP2, ~QSB, des.wt, svymean, keep.var=TRUE, na.rm=T)
soul$QSB2 <- soul$QSB 
soul <- merge(soul,att.pr[,1:2], by='QSB')
soul$QSB <- with(soul, reorder(QSB2, CCEGRP2Attached)  ) 
soul$CCEGRP2Attached <- NULL

@@

# select variables with less than 25 % of missing values 
var.red <- names(aux[aux < .25])
aux     <- apply(miss[,var.red], 1, sum )
aux1    <- data.frame(table(aux))
soul.red <- soul[aux <3, var.red]
orig.red <- colnames(soul.red)[c(4,5,7,8:50)]

# age and proportion of age in place where live. 
soul <- mutate(soul,QD1gr=cut(QD1, breaks=c(0,30,40,50,60,70,80,99), include.lowest=T),
               QD2gr=cut(QD2, breaks=c(0,10,20,50  ,99), include.lowest=T),prop.res =QD2/QD1,
               prop.gr=cut(prop.res, breaks=c(0,0.25,0.5,1,5.5),include.lowest=T))

soul$prop.res[soul$QD1==0] <- NA

demo.red <- c(colnames(soul.red)[c(6,51:62)],'QD1gr','QD2gr','prop.res','prop.gr')

@@

# QD10 ask for hispanic or not hispanic, QD111 ask for race, we combine both in QD111                      
aux <- reco.demo$QD111
levels(aux)[c(4,6:10,15:17,19,20) ] <- 'Other'
levels(aux)[c(1,2,5,7,8,10) ] <- 'Non Response'
levels(aux)[2] <- 'Hispanic'
aux[reco.demo$QD10=='Yes'] <- 'Hispanic'                      
reco.demo$QD111  <- factor(aux)

@@

<<census,dependson='data', echo=FALSE,results='hide',warning=FALSE,message=FALSE,eval=FALSE>>=

# 1) read data tables from census

file <-c( "agesex.csv","education.csv","income.csv","owner.csv","race.csv","workers.csv", "yearentry.csv")

for (i in 1:length(file) )  {
  aux <- read.csv(file[i], header=T)
  aux <- subset(aux, X=='Estimate')

  for (j in 4:dim(aux)[2]) aux[,j] <- as.numeric(as.character(aux[,j]))

  aux1 <- subset(aux,select=-c(1,3),subset=!community%in%c('Gary','Philadelphia'))
  aux2 <- apply(subset(aux,community=='Gary',select=-c(1,2,3)),2,sum)
  aux2 <- data.frame(community='Gary',t(aux2))

  aux3 <- apply(subset(aux,community=='Philadelphia',select=-c(1,2,3)),2,sum)
  aux3 <- data.frame(community='Philadelphia',t(aux3))

aux <- rbind(aux1,aux2,aux3)
 aux$community <- factor(aux$community)
  nam <- sub(pattern='.csv', x=file[i], replacement='')
  assign(nam, aux)
}
# for incomePC is slightly different ... 
aux <- read.csv("incomePC.csv", header=T)
aux <- subset(aux, X=='Estimate')
aux[,4] <- as.numeric(as.character(aux[,4]))
  aux1 <- subset(aux,select=-c(1,3),subset=!community%in%c('Gary','Philadelphia'))
  aux2 <- apply(subset(aux,community=='Gary',select=-c(1,2,3)),2,mean)
  aux2 <- data.frame(community='Gary',t(aux2))

  aux3 <- apply(subset(aux,community=='Philadelphia',select=-c(1,2,3)),2,mean)
  aux3 <- data.frame(community='Philadelphia',t(aux3))

aux <- rbind(aux1,aux2,aux3)
 aux$community <- factor(aux$community)
  assign('incomePC', aux)

remove(aux, aux1, aux2,aux3, i, j, nam)

# 2) create a data set with variables from every, topic. We compute some variables per topic
# and save it in census.data
census.data <- agesex[,1]

# agesex    has population by sex and age, the columns ending with 'F' are for female. 
# for now : agesex[,c(2,4,5,29)],  
x <- which(colnames(agesex) == 'Female.')
colnames(agesex)[-c(1:x)] <- paste(colnames(agesex)[-c(1:x)],'F', sep='')
p <- c(2.5, 7,12,16,18.5,20,21,23,27,32,37,42,47,52,57,60.5,63,65.5,68,72,77,82,90)
pp <- matrix(p, nrow=26, ncol=23, byrow=T)

agesex <- mutate(agesex,aveageM = apply(agesex[,4:26]*pp, 1, sum)/agesex$Male,
                 aveageF= apply(agesex[,28:50]*pp, 1, sum)/agesex$Female,
                 prop.male=agesex$Male/agesex$Total,
                 prop.female=agesex$Female/agesex$Total)

agesex <- mutate(agesex,prop.m.f = agesex$aveageF/agesex$aveageM,
                 pop100=agesex[,2]/100000,                
                 prop1821=apply(agesex[,c(8:10,32:34)], 1, sum)/agesex$Total)

census.data <- cbind(census.data,agesex[,c(51:57)])

# -----------------------------------------
# education has info, about education level and working status, unemployment rate can be 
# computed from this. Universe: Pop between 25-64 years
# un rate 
education <- mutate(education,unemployed = apply(education[, c('Unemployed','Unemployed.1','Unemployed.2','Unemployed.3')]
                                          ,1,sum),
                              employed = apply(education[, c('Employed','Employed.1','Employed.2','Employed.3')],1,sum),
                              militar = apply(education[, c('In.Armed.Forces','In.Armed.Forces.1','In.Armed.Forces.2',
                                                            'In.Armed.Forces.3')],1,sum),
                              not.labor = apply(education[, c('Not.in.labor.force','Not.in.labor.force.1',
                                                              'Not.in.labor.force.2','Not.in.labor.force.3')], 1, sum))

aux2 <- apply(education[, c('In.labor.force.','In.labor.force..1','In.labor.force..2','In.labor.force..3')], 1, sum)

education <- mutate(education,un.rate =unemployed/aux2,pr.rate = aux2/education$Total.)

pet <- apply(agesex[, c(12:20,36:44)],1,sum)

census.data <- cbind(census.data, education[,c(3,10,17,24,31:34)]/education$Total., education[,c(35,36)] )

#-------------------------------------
# income    has info about income level, we should pooled some categories, or some meassure of inequality can be computed here and compute average income per community.

# mean point for income levels in miles (5=5000), we are not sure about the 
# upper limit, range is "over 200". 
p2 <- c(5, 12.5,17,22.5,27,32.5,37.5,42.5,47.5,55,67.5,87.5,112.5,137.5,175,200)
pp2 <- matrix(p2, nrow=26, ncol=16, byrow=T)

income <- mutate(income,av.income=(apply(income[,-c(1,2)]*pp2,1,sum)/income$Total.)*1000,
                 income.pc=(apply(income[,-c(1,2)]*pp2,1,sum)/agesex$Total.)*1000)

income$gini <- 0
for (i in 1:26) {
  d <- rbind(mid=1000*p2, cont=as.vector(income[i,3:18]) ) 
  d <- data.frame(t(d))
  income$gini[i]  <- with(d, gini(mid, weights = cont  ) ) 
}
census.data <- cbind(census.data, income[,19:21])

#-------------------------------------------------------
# owner: People living in housholds, renting vs owning and how long are living in the same place.

colnames(owner)[2] <- 'Total'
p3 <- c(5,10,15,25,35,45)
pp <- matrix(rep(p3,2), nrow=26, ncol=6, byrow=T)

owner <- mutate(owner,own.rate = Owner.occupied./Renter.occupied.,
                ow.prop10 = apply(owner[,6:9],1,sum)/owner$Owner.occupied.,
                ren.prop10 = apply(owner[,13:16],1,sum)/owner$Renter.occupied.,
                av.rent = apply(owner[,c(11:16)]*pp,1,sum)/owner$Renter.occupied.,
                av.own = apply(owner[,c(4:9)]*pp,1,sum)/owner$Owner.occupied.)

census.data <- cbind(census.data, owner[,17:21])

#---------------------------------------------------------
# race, we can use proportion os white people, and also entropy measure with the 
# races distribution in each community, sum(p*log(p))

race <- mutate(race,white.prop = White/Total.,afri.prop = Black/Total., asian.prop = Asian/Total.)
aux <- apply(race[,-c(1,2,3,4,6)],1,sum)
race$other.prop <- aux/race$Total.

census.data <- cbind(census.data, race[,12:15])

#-----------------------------------------------------------
# workers
#Proportion of household with 1,2 or more than 3 workers (prop.1worker,prop.2worker,prop.3worker)
#Proportion of houses with one two ando so on persons (prop.1.per,prop.2.per,prop.3.per,prop.3.per)
#Proportion of no workers and one worker in  households with only one person  (prop1.no.workers, prop1.worker)
#proportion of no workers, one worker and two workers in  households with two person (prop2.no.workers, prop2.1worker,prop.2worker )
#same than previous but for households with Three person househole (prop3.no.workers,prop3.1worker, prop3.2worker,prop3.3worker  )
#same than previous but for households with four person or more househole (prop3.no.workers,prop3.1worker,prop3.2worker,prop3.3worker)

workers <- mutate(workers,prop.no.workers = workers$No.workers/workers$Total.,
                  prop.1worker = workers$X1.worker/workers$Total., 
                  prop.2worker = workers$X2.workers/workers$Total., 
                  prop.3worker = workers$X3.or.more.workers/workers$Total.,
                  prop.1.per = workers$X1.person.household./workers$Total., 
                  prop.2.per = workers$X2.person.household./workers$Total., 
                  prop.3.per = workers$X3.person.household./workers$Total.,
                  prop.4.per= workers$X4.or.more.person.household./workers$Total,
                  prop1.no.workers= workers$No.workers.1/workers$X1.person.household.,
                  prop1.worker = workers$X1.worker.1/workers$X1.person.household.,
                  prop2.no.workers =  workers$No.workers.2/workers$X2.person.household.,
                  prop2.1worker = workers$X1.worker.2/workers$X2.person.household.,
                  prop2.2worker = workers$X2.workers.1/workers$X2.person.household.,
                  prop3.no.workers = workers$No.workers.2/workers$X3.person.household.,
                  prop3.1worker = workers$X1.worker.3/workers$X3.person.household.,
                  prop3.2worker = workers$X2.workers.2/workers$X3.person.household.,
                  prop3.3worker = workers$X3.workers/workers$X3.person.household.,
                  prop3.no.workers = workers$No.workers.2/workers$X3.person.household.,
                  prop3.1worker = workers$X1.worker.3/workers$X3.person.household.,
                  prop3.2worker = workers$X2.workers.2/workers$X3.person.household.,
                  prop3.3worker = workers$X3.workers/workers$X3.person.household.)

census.data <- cbind(census.data, workers[,24:40])
#-----------------------------------------------------------------------
# yearentry 
#proportion of person entry 2000 or later (prop.new.entry,prop.entry.90s,prop.entry.80s,prop.entry.prev.80s)
#proportiono of native and foreign for different entries (prop.nat.new.entry,prop.for.new.entry,prop.nat.entry.90s,prop.for.entry.90s)
#proportion of persons not original from the place (entred,entred.nat, entred.for)
#we can separate foreign born in naturalized us citizen and not us citizen (entred.for.nat.us, entred.for.not.us)

yearentry <- mutate(yearentry,prop.new.entry = yearentry$Entered.2000.or.later/yearentry$Total.,
                    prop.entry.90s = yearentry$Entered.1990.to.1999./yearentry$Total.,
                    prop.entry.80s = yearentry$Entered.1980.to.1989./yearentry$Total.,
                    prop.entry.prev.80s = yearentry$Entered.before.1980./yearentry$Total.,
                    prop.nat.new.entry = yearentry$Native/yearentry$Entered.2000.or.later,
                    prop.for.new.entry = yearentry$Foreign.born./yearentry$Entered.2000.or.later,
                    prop.nat.entry.90s = yearentry$Native.1/yearentry$Entered.1990.to.1999.,
                    prop.for.entry.90s = yearentry$Foreign.born..1/yearentry$Entered.1990.to.1999.,
                    prop.nat.entry.80s = yearentry$Native.2/yearentry$Entered.1980.to.1989.,
                    prop.for.entry.80s = yearentry$Foreign.born..2/yearentry$Entered.1980.to.1989.,
                    prop.nat.entry.prev.80s = yearentry$Native.3/yearentry$Entered.before.1980.,
                    prop.for.entry.prev.80s = yearentry$Foreign.born..3/yearentry$Entered.before.1980.,
                    entred = yearentry$Total./agesex$Total.,
                    entred.nat = yearentry$Native/agesex$Total.,
                    entred.for = yearentry$Foreign.born./agesex$Total.,
                    entred.for.nat.us = yearentry$Naturalized.U.S..citizen/agesex$Total.,
                    entred.for.not.us = yearentry$Not.a.U.S..citizen/agesex$Total.)

census.data <- cbind(census.data, yearentry[,23:38])
census.data$income.pc <- incomePC$income.pc
colnames(census.data)[1] <- 'QSB'

# Compute comunity proportion of attached people, carefull casuse comunities are 
# in different order in soul data and census data. 
aux <- with(soul,table(QSB,CCEGRP2))
tot <- apply(aux,1,sum)

prop.atach <- data.frame(aux[,1]/tot,aux[,2]/tot,aux[,3]/tot)
colnames(prop.atach) <- colnames(aux)
prop.atach <- prop.atach[order(rownames(prop.atach)),]
census.data <- census.data[order(census.data[,1]), ]

census.data <- data.frame(QSB=rownames(prop.atach), census.data[,-1],prop.atach) 
census.data$QSB <- with(census.data, reorder(QSB, Attached)  )

#-----------------------------
corrs <- cor(census.data[,-1])
aux1 <- sort(corrs[,'Attached'] )
dat.cor <- rbind(data.frame(vars = names(corrs[1:61,'Attached']), cor = corrs[1:61,'Attached'] ,
                            positive=corrs[1:61,'Attached'] > 0, type='Attached') , data.frame(vars = names(corrs[1:61,'Not.Attached']), 
                            cor = corrs[1:61,'Not.Attached'] , positive=corrs[1:61,'Not.Attached'] > 0,type='Not.Attached'))

rownames(dat.cor) <- NULL
dat.cor$vars <- reorder(dat.cor$vars, rep(dat.cor$cor[dat.cor$type=='Attached'],2) )   
dat.cor2 <- subset(dat.cor, vars %in% names(aux1[c(2:11, 51:61)]) )
dat.cor2$vars <- factor(dat.cor2$vars)
@

<<cluster, dependson='data',echo=FALSE,warning=FALSE, cache=TRUE,eval=FALSE>>=
# we choose 4 clusters
  set.seed(1234); km <- kmeans(soul.red2[,c(3,9:45)],centers=4)

# relation with attach, one cluster attach, one cluster not attach, 
# two clusters in the middle 
soul.red2$cl.id <- factor(km$cluster)

# Characterize the clusters with the numerical variables
  pr <- soul.red2$PROJWT/3 
 des.pro <- svydesign(ids=~1, weights=~pr, data= soul.red2)

frm <- formula(paste('~',paste(names(soul.red2[,3:45]), collapse='+')))

means <-svyby(frm, ~cl.id+ YEAR, data=soul.red2,design=des.pro, FUN=svymean, keep.var=TRUE, na.rm=T)
means <- means[,-grep('se.', colnames(means))]

  des.cl1  <- melt(means, id.vars=c('cl.id','YEAR'))
  cv <- function(x) min(x) -max(x) 
  des.cl1$variable <- reorder(des.cl1$variable, des.cl1$value, cv )

@

<<importance, dependson=c('data','cluster'), echo=FALSE,results='hide', cache=TRUE,warning=FALSE,message=FALSE,eval=FALSE>>=
# Fit RF on each community, to explain CCE with the constructed variables

fr <- as.formula('CCEGRP2~. -CCE-cl.id-QS3_02-Q3A-Q3B-Q3C-QCE1-QCE2-QS3-WEIGHT-PROJWT')
bosque <- dlply(soul.red2, .(QSB), function(x) randomForest(fr,importance=T, data=x))
xx <- ldply(bosque, function(x) t(importance(x)[,'MeanDecreaseGini']))
xx <- data.frame(QSB=names(bosque), xx)

important <- melt(xx, id.vars='QSB')
colnames(important) <- c('QSB', 'dim', 'mdg')
important <- ddply(important, .(QSB), mutate, mdg.pr = 100*mdg/sum(mdg), rnk = rank(-mdg) )

mdg.mm = ddply(important, .(dim), summarise, mdg.mm = mean(mdg.pr) )

important <- merge(important,mdg.mm)
important$rmdg = with(important, (mdg.pr-mdg.mm) *(mdg.pr > mdg.mm))

rankings <- ddply(subset(important, rnk < 6), .(dim,rnk), function(x) length(x$dim) )
rankings$rnk <- as.factor(rankings$rnk)
levels(rankings$rnk) <- paste('R', 1:5, sep='')

# reorder QSB levels again
important <- merge(important,att.pr[,1:2], by='QSB')
important$QSB <- with(important, reorder(QSB, CCEGRP2Attached)  ) 
important$CCEGRP2Attached <- NULL

@

@@

br <-  with(soul,quantile(OPENNESS,prob=seq(0,1, .2), na.rm=T)) 
soul$Openess <- with(soul, cut(OPENNESS,br))
levels(soul$Openess) <- round(br,1)

br <-  with(soul,quantile(SOCIAL_O,prob=seq(0,1, .2), na.rm=T)) 
soul$Social.Offering <- with(soul, cut(SOCIAL_O,br))
levels(soul$Social.Offering) <- round(br,1)

prodplot(data=soul,PROJWT/3 ~CCEGRP2+ Openess, divider=c("vspine", "hspine"),subset=level == 2) +aes(fill=CCEGRP2) + theme(legend.position = 'none',axis.title.y=element_blank()) + scale_fill_brewer(type='seq', palette=1)

prodplot(data=soul, PROJWT/3 ~ CCEGRP2+ Social.Offering , divider=c("vspine", "hspine"),subset=level == 2)+aes(fill=CCEGRP2) + theme(axis.text.y = element_blank(),axis.title.y=element_blank()) + scale_fill_brewer(type='seq', palette=1, name='Attach level')

@@

The Knight survey data contains missing information. The proportion of missing values was computed for each observation (by cases) and for each variable (by variables). Table \ref{tab1} shows the quantiles of the proportion of missing values computed by case and computed by variable. Focusing on the proportions of missing values by case, there is 19\% of missing record for the most complete response. This may be explained by the fact that 3 years of data was combined although some variables only appeared in 1 or 2 years. On the other hand, focusing on the proportion of missing values by variables, most of the variables show a reasonable proportion of missing values (13\%) while there are a few variables with really high proportions of missing values.

@@

miss.c2 <- apply(miss[, miss.v < .25] , 1, mean) 
tab2 <- rbind(quantile(miss.v),quantile(miss.c) )
rownames(tab2) <- c('by Variable','by Case')
colnames(tab2) <- c('Min', 'Q25', 'Median', 'Q75', 'Max')
cap <- 'Quantiles of the proportion of missing values by case and variables.'
xt <- xtable(tab2,caption=cap,label='tab1', align= 'lccccc' )
print(xt, ,caption.placement="top")

complete <- 100*sum(miss.c2==0)/length(miss.c2)
dim.case <- dim(soul)[1]*sum(miss.c2==0)/length(miss.c2)
dim.var <- sum(miss.v < .25)
@

@@

Figure \ref{map} shows the communities location, most are from the East Coast of the United States. Size of the dot represents population. Detroit and Philadelphia are the biggest communities while Aberdeen and Milledgeville are the smallest.
\begin{figure}[hbpt] 
\centering
<<mapita,dependson='data',echo=FALSE, fig.height=6,message=FALSE,warning=FALSE,warning=FALSE>>=
lon.lat <- read.table('lon.lat.txt',header=T)
require(ggmap)
require(mapproj)

@@

aux.mer <- merge(lon.lat,census.data[,c(1,7)],by.x='Community',by.y='QSB')
aux.mer$cate[aux.mer$pop100<3] <- 1
aux.mer$cate[aux.mer$pop100>=3&aux.mer$pop100<8] <- 2
aux.mer$cate[aux.mer$pop100>=8&aux.mer$pop100<20] <- 3
aux.mer$cate[aux.mer$pop100>=20] <- 4

p <- ggmap(map, extent = "device")
p+geom_point(data=aux.mer, mapping=aes(lon, lat),size=6*aux.mer$cate,alpha=.4)+xlab('')+ylab('')+theme(axis.text.x=element_blank(),axis.text.y=element_blank()) 
@
  \caption{Location of the 26 communities analyzed, size of dot represent population. Only two communities on the west coast. \label{map}}
\end{figure}

@@

dat.aux<-melt(data=agesex,id.vars=c('community','prop1821'),measure.vars=c(4:26,28:50) )
dat.aux$Gender <- rep( c('Male','Female'), each=23*26)
levels(dat.aux$variable)[24:46] <- levels(dat.aux$variable)[1:23] 
dat.aux$prop <- with(dat.aux, value/rep(agesex$Total., times=23*2))

p <- c(2.5, 7,12,16,18.5,20,21,23,27,32,37,42,47,52,57,60.5,63,65.5,68,72,77,82,90)
levels(dat.aux$variable) <- p
dat.aux$variable <- as.numeric(as.character(dat.aux$variable))
dat.aux$community <- with(dat.aux, reorder(community, -prop1821) )

qplot(data=dat.aux,x=variable,y=prop,colour=Gender,shape=Gender)+xlab("Age")+ylab("Proportion")+geom_line( ) +geom_point(size=I(.1))+ facet_wrap(facets=~community,ncol=7)+ theme(legend.position=c(.9,0.1)) + scale_color_brewer(type='qual', palette='Dark2')

@@

\begin{figure}
<<education,echo=FALSE, dependson='data'>>=
dat.aux <- melt(data=census.data,id.vars='QSB',measure.vars=c("Less.than.high.school.graduate.","High.school.graduate.","Some.college.or.associate.s.degree.","Bachelor.s.degree.or.higher.") )

or.aux <- rep(dat.aux$value[dat.aux$variable=="Bachelor.s.degree.or.higher."],4)
dat.aux$QSB <- with(dat.aux, reorder(x=QSB, X=or.aux  ) )

@@

prodplot(data=dat.aux, value~Education.Level+QSB,divider=c( "hspine","vspine"))+aes(fill=Education.Level) + theme(legend.position="bottom",legend.title=element_blank(),axis.title.y=element_blank(),axis.text.y = element_text( size=I(10)) ) + scale_fill_brewer(type='seq', palette=2)
@
\caption{Product plot of education level by community, the width of each box represent the proportion of people with that education level within a community.  Communities are sorted from top to bottom by proportion of people with college degree. Big range of variation on education level among the communities. \label{edu}}
\end{figure}

Figure \ref{edu} shows the proportion of people in each education level across communities, sorted from top to bottom by proportion of people with a college degree. Since the data from the \cite{acs2011} is already aggregated at the community level, this product plot only reports proportion of education within each community (not joint proportions), which is why all horizontal bars have the same height. 
We can observe that Boulder, San Jose, and State College are the most educated communities while Milledgeville, Biloxi, and Macon are the less educated communities.
\begin{figure}
<<income, echo=FALSE,dependson=c('data','census'), fig.height=6,fig.keep='all', out.width='0.49\\linewidth'>>=
aux <- subset(census.data, QSB %in% levels(census.data$QSB)[c(18,9,20,16,5,3,10,1,7,8,14)] )
aux <- mutate(aux, gr =factor(c(2,1,2,2,3,3,3,1,1,3,2)), h = c(1,0,0,0,1,1,1,0,0,0,1), v = c(0,0,0,1,0,1,0,0,0,0,0))

>>>>

% All chunks that make models or data procesing, at the begining
<<setup, include=FALSE, cache=FALSE,warning=FALSE,message=FALSE>>=
options(replace.assign=TRUE,width=50)
opts_chunk$set(fig.path='figure/graphics-', cache.path='cache/graphics-', fig.align='center', fig.width=8, fig.height=5, fig.show='hold', cache=TRUE, par=TRUE)
knit_hooks$set(par=function(before, options, envir){
if (before && options$fig.show!= 'none') par(mar = c(4, 4, .1, .1), cex.lab = .95, cex.axis = .9, mgp = c(2, .7, 0), tcl = -.3)
}, crop = hook_pdfcrop)

# packages
library(reshape2)
library(plyr)
library(productplots) 
library(ggplot2)
library(stats)
library(randomForest)
library(dplyr)
library(tidyr)
library(maps)
library(fpc)
library(xtable)
library(car)
library(GGally)
library(survey)
library(reldist)

@@

# Compute comunity proportion of attached people
lev <- levels(soul$QSB)
census.data %<>%
        mutate(QSB = factor(census.data$QSB, labels=sort(lev)) ) %>%
        left_join(att.pr[, c(1,2,4)], by = "QSB") %>% 
        mutate(QSB = reorder(QSB, CCEGRP2Attached))

@@

# Variable sets 
orig <- c(4:6, 8:61 ,148:152, 165:175, 181:206, 210:213)
demo <- c(7, 62:79, 153:154, 158:160, 206:209, 216, 229)
reco <- c(80:120, 141:147, 161, 177:179, 217:228)
cons <- c(121:135, 155:157, 162:164, 176, 180)

@@

# Compute attachment proportion for ordering QSB levels
wt     <- with(subset(soul,!is.na(WEIGHT)), WEIGHT/3)
des.wt <- svydesign(ids=~1, strata = ~QSB, weights = ~wt, data = subset(soul, !is.na(WEIGHT) ) )
att.pr <- svyby(~CCEGRP2, ~QSB, des.wt, svymean, keep.var = TRUE, na.rm = T)
soul$QSB2 <- soul$QSB 
soul <- merge(soul,att.pr[,1:2], by = 'QSB')
soul$QSB <- with(soul, reorder(QSB2, CCEGRP2Attached)  ) 
soul$CCEGRP2Attached <- NULL

@@

# select variables with less than 25 % of missing values 
var.red <- names(aux[aux < .25])
aux     <- apply(miss[,var.red], 1, sum )
aux1    <- data.frame(table(aux))
soul.red <- soul[aux < 3, var.red]
orig.red <- colnames(soul.red)[c(4, 5, 7, 8:50)]

# age and proportion of age in place where live. 
soul <- mutate(soul, QD1gr = cut(QD1, breaks = c(0, 30, 40, 50, 60, 70, 80, 99), include.lowest = T),
               QD2gr=cut(QD2, breaks = c(0, 10, 20, 50, 99), include.lowest = T), prop.res = QD2/QD1,
               prop.gr=cut(prop.res, breaks = c(0, 0.25, 0.5, 1, 5.5), include.lowest = T))

soul$prop.res[soul$QD1 == 0] <- NA

demo.red <- c(colnames(soul.red)[c(6, 51:62)],'QD1gr','QD2gr','prop.res','prop.gr')

@@

# QD10 ask for hispanic or not hispanic, QD111 ask for race, we combine both in QD111                      
aux <- reco.demo$QD111
levels(aux)[c(4, 6:10, 15:17, 19, 20) ] <- 'Other'
levels(aux)[c(1, 2, 5, 7, 8, 10) ] <- 'Non Response'
levels(aux)[2] <- 'Hispanic'
aux[reco.demo$QD10=='Yes'] <- 'Hispanic'                      
reco.demo$QD111  <- factor(aux)

@@

@@

br <-  with(soul, quantile(OPENNESS, prob = seq(0,1, .2), na.rm = T)) 
soul$Openess <- with(soul, cut(OPENNESS, br))
levels(soul$Openess) <- round(br, 1)

br <-  with(soul,quantile(SOCIAL_O, prob = seq(0, 1, .2), na.rm = T)) 
soul$Social.Offering <- with(soul, cut(SOCIAL_O, br))
levels(soul$Social.Offering) <- round(br, 1)

prodplot(data = soul, PROJWT/3 ~CCEGRP2 + Openess, divider = c("vspine", "hspine"), subset = level == 2) + aes(fill = CCEGRP2) + theme(legend.position = 'none',axis.title.y = element_blank()) + scale_fill_brewer(type = 'seq', palette = 1)

prodplot(data = soul, PROJWT/3 ~ CCEGRP2 + Social.Offering , divider = c("vspine", "hspine"), subset = level == 2) + aes(fill = CCEGRP2) + theme(axis.text.y = element_blank(),axis.title.y = element_blank()) + scale_fill_brewer(type = 'seq', palette = 1, name = 'Attach level')

@@

The Knight survey data contains missing information. The proportion of missing values was computed for each observation (by cases) and for each variable (by variables). Table \ref{tab1} shows the quantiles of the proportion of missing values computed by case and computed by variable. Focusing on the proportions of missing values by case, there is 19\% of missing record for the most complete response. This may be explained by the fact that 3 years of data was combined although some variables only appeared in 1 or 2 years. On the other hand, focusing on the proportion of missing values by variables, most of the variables show a reasonable proportion of missing values (the median is 12\%) while there are a few variables with really high proportions of missing values.

@@

miss.c2 <- apply(miss[, miss.v < .25], 1, mean) 
tab2 <- rbind(quantile(miss.v), quantile(miss.c) )
rownames(tab2) <- c('by Variable', 'by Case')
colnames(tab2) <- c('Min', 'Q25', 'Median', 'Q75', 'Max')
cap <- 'Quantiles of the proportion of missing values by case and variables.'
xt <- xtable(tab2, caption = cap, label = 'tab1', align = 'lccccc' )
print(xt, ,caption.placement = "top")

complete <- 100*sum(miss.c2 == 0)/length(miss.c2)
dim.case <- dim(soul)[1]*sum(miss.c2 == 0)/length(miss.c2)
dim.var <- sum(miss.v < .25)
@

@@

Figure \ref{map} shows the communities location, most are from the East Coast of the United States. Size of the dot represents population. Detroit and Philadelphia are the biggest communities while Aberdeen and Milledgeville are the smallest.
\begin{figure}[hbpt] 
\centering
<<mapita,dependson='data',echo=FALSE, fig.height=6,message=FALSE,warning=FALSE,warning=FALSE, cache=TRUE>>=
lon.lat <- read.table('lon.lat.txt',header=T)
library(ggmap)
library(mapproj)

@@

aux.mer <- merge(lon.lat, census.data[, c(1, 7)], by.x = 'Community', by.y = 'QSB')
aux.mer$cate[aux.mer$pop100 < 3] <- 1
aux.mer$cate[aux.mer$pop100 >= 3&aux.mer$pop100 < 8] <- 2
aux.mer$cate[aux.mer$pop100 >= 8 & aux.mer$pop100 < 20] <- 3
aux.mer$cate[aux.mer$pop100 >= 20] <- 4

p <- ggmap(map, extent = "device")
p + geom_point(data = aux.mer, mapping = aes(lon, lat), size = 6*aux.mer$cate, alpha = .4) + xlab('') + ylab('') + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) 
@
  \caption{Location of the 26 communities analyzed, size of dot represent population. Only two communities on the west coast. \label{map}}
\end{figure}

@@

dat.aux <- melt(data = agesex, id.vars = c('community', 'prop1821'), measure.vars = c(4:26, 28:50))
dat.aux$Gender <- rep( c('Male','Female'), each = 23*26)
levels(dat.aux$variable)[24:46] <- levels(dat.aux$variable)[1:23] 
dat.aux$prop <- with(dat.aux, value/rep(agesex$Total., times=23*2))

p <- c(2.5, 7, 12, 16, 18.5, 20, 21, 23, 27, 32, 37, 42, 47, 52, 57, 60.5, 63, 65.5, 68, 72, 77, 82, 90)
levels(dat.aux$variable) <- p
dat.aux$variable <- as.numeric(as.character(dat.aux$variable))
dat.aux$community <- with(dat.aux, reorder(community, -prop1821) )

qplot(data = dat.aux, x = variable,y = prop, colour = Gender, shape = Gender) + xlab("Age") + ylab("Proportion") + geom_line( ) + geom_point(size = I(.1)) + facet_wrap(facets =~community, ncol = 7) + theme(legend.position = c(.9,0.1)) + scale_color_brewer(type = 'qual', palette = 'Dark2')

@@

\begin{figure}
<<education,echo=FALSE, dependson='data'>>=
dat.aux <- census.data %>% 
  select(QSB, Less.than.high.school.graduate.,High.school.graduate.,Some.college.or.associate.s.degree.,Bachelor.s.degree.or.higher.) %>% arrange(Bachelor.s.degree.or.higher.) %>%
  mutate(QSB = reorder(QSB, Bachelor.s.degree.or.higher.)) %>%
  gather(variable, value, -QSB)

#dat.aux <- melt(data=census.data,id.vars='QSB',measure.vars=c("Less.than.high.school.graduate.","High.school.graduate.","Some.college.or.associate.s.degree.","Bachelor.s.degree.or.higher.") ) 
#or.aux <- rep(dat.aux$value[dat.aux$variable=="Bachelor.s.degree.or.higher."],4)
#dat.aux$QSB <- with(dat.aux, reorder(x=QSB, X=or.aux  ) )

@@

prodplot(data = dat.aux, value~Education.Level + QSB, divider = c( "hspine","vspine")) + aes(fill = Education.Level) + theme(legend.position = "bottom", legend.title = element_blank(), axis.title.y = element_blank(), axis.text.y = element_text( size = I(10)) ) + scale_fill_brewer(type = 'seq', palette = 2)
@
\caption{Product plot of education level by community, the width of each box represent the proportion of people with that education level within a community.  Communities are sorted from top to bottom by proportion of people with college degree. Big range of variation on education level among the communities. \label{edu}}
\end{figure}

Figure \ref{edu} shows the proportion of people in each education level across communities, sorted from top to bottom by proportion of people with a college degree. Since the data from the \cite{acs2011} is already aggregated at the community level, this product plot only reports proportion of education within each community (not joint proportions), which is why all horizontal bars have the same height. 
We can observe that Boulder, San Jose, and State College are the most educated communities while Milledgeville, Biloxi, and Macon are the less educated communities.
\begin{figure}
<<income, echo=FALSE,dependson=c('data'), fig.height=6,fig.keep='all', out.width='0.49\\linewidth'>>=

grps <- c("Boulder, CO", "San Jose, CA" ,"St. Paul, MN",
"Aberdeen, SD", "Wichita, KS", "Fort Wayne, IN", "Gary, IN",
"Macon, GA", "Miami, FL", "Milledgeville, GA", "Tallahassee, FL")
